\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}

\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue}

\lstset{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue},
  commentstyle=\color{gray},
  frame=single,
  breaklines=true,
  columns=fullflexible,
  language=Verilog
}

\title{FPGA Session Log --- \texttt{final\_8}: 4-to-16 Decoder\\
\large Digilent Arty A7-35T \,\textperiodcentered\, Open-Source Toolchain}
\author{Lucas Zhou}
\date{July 13, 2026}

\begin{document}
\maketitle

\section{Session Goal}
Compile the hand-written \texttt{top.v} (a 4-to-16 decoder mapping two 2-bit
inputs $\{A2,A1\}$ and $\{B2,B1\}$ to a seven-segment pattern
\texttt{seg[6:0]}), create a testbench, and establish the simulation flow.
The Pmod SSD (seven-segment display) hardware is on order.

\section{Compilation Result}
Synthesis was attempted with Yosys 0.66 (oss-cad-suite):
\begin{lstlisting}[language=bash]
yosys -p "synth_xilinx -flatten -abc9 -arch xc7 -top top; \
          write_json top.json" top.v
\end{lstlisting}
It failed at the first parse error:
\begin{lstlisting}[language={}]
top.v:2: ERROR: syntax error, unexpected TOK_ID,
         expecting ')' or ',' or '='
\end{lstlisting}

Three recurring syntax issues were identified (fixes to be applied by the
author):
\begin{enumerate}
  \item \textbf{Port list commas} (line 2): ports must be comma-separated,
        including between the \texttt{input} and \texttt{output}
        declarations, e.g.\ \texttt{input wire A2, A1, B2, B1,}.
  \item \textbf{Case-item colons} (lines 9--25): each case item requires a
        colon, e.g.\ \texttt{4'b0000: seg = ...} (likewise
        \texttt{default:}).
  \item \textbf{Statement semicolons}: every assignment
        \texttt{seg = 7'b...} must end with \texttt{;}.
\end{enumerate}
The decoder logic itself (the 2-bit $\times$ 2-bit product table) is expected
to compile cleanly once these are corrected.

\section{Testbench}
\texttt{tb\_top.v} was created. It sweeps all 16 combinations of
$\{A2,A1,B2,B1\}$, prints the resulting \texttt{seg[6:0]} pattern for each,
and dumps \texttt{tb\_top.vcd} for waveform viewing. It is deliberately
\emph{not} self-checking: the printed patterns are to be compared by hand
against the case table in \texttt{top.v}.

\section{Simulation Flow}
From \texttt{projects/final\_8/}:
\begin{lstlisting}[language=bash]
source /Users/lucaszhou/15A/FPGAproj1/tools/oss-cad-suite/environment
iverilog -g2005 -o tb_top.vvp tb_top.v top.v   # compile
vvp tb_top.vvp                                 # run, prints 16 cases
gtkwave tb_top.vcd                             # optional waveforms
\end{lstlisting}

\section{Follow-up Session (same day): XDC Rewrite}

The author fixed the case-item colons and statement semicolons in
\texttt{top.v}; one syntax issue remains (port-list commas, line~2).
\texttt{arty.xdc} was then rewritten to match the current ports:

\begin{itemize}
  \item \textbf{Inputs}: \texttt{A2/A1/B2/B1} on the slide switches,
        keeping the earlier assignment A2=SW3 (A10), A1=SW2 (C10),
        B2=SW1 (C11), B1=SW0 (A8).
  \item \textbf{Outputs}: \texttt{seg[6:0]} on Pmod connector \textbf{JA}
        (top-right of the board), where the Pmod SSD's segment inputs land:
        \texttt{seg[0]}$\to$JA1 (G13, signal AA) through
        \texttt{seg[6]}$\to$JA9 (A18, signal AG); i.e.\ \texttt{seg[N]}
        drives SSD segment $N{+}1$. If a segment lights in the wrong
        position on real hardware, permute the XDC pin assignments rather
        than the \texttt{top.v} case table.
  \item \textbf{Digit select}: the Pmod SSD is a two-digit display; JA
        pin~10 (K16, signal C) selects the lit digit. The constraint is
        present but commented out because \texttt{top.v} has no such port
        yet.
\end{itemize}

\textbf{Verification}: a scratch copy of \texttt{top.v} (with only the
line-2 commas added; the author's file was not modified) was run through
Yosys and nextpnr-xilinx with the new XDC --- clean synthesis, all 11
ports constrained to their pads, and a FASM file generated. The real
build should therefore pass once the author applies the comma fix.

\subsection*{Adding the digit-select port (to be applied by the author)}
\begin{lstlisting}
module top (
  input wire A2, A1, B2, B1, //bits
  output reg [6:0] seg, //from 1 to 7
  output wire C //digit select for Pmod SSD
);
...
assign C = 1'b1; // outside the always block; flip to 1'b0 if the wrong digit lights
\end{lstlisting}
After adding the port, uncomment the K16 line at the bottom of
\texttt{arty.xdc}.

\section{Open Items}
\begin{itemize}
  \item \textbf{Fix \texttt{top.v} port commas} (author) --- line 2:
        \texttt{input wire A2, A1, B2, B1,}.
  \item \textbf{Add digit-select port \texttt{C}} (author) --- snippet
        above; then uncomment the K16 constraint in \texttt{arty.xdc}.
  \item \textbf{Confirm on hardware once the Pmod SSD arrives} ---
        segment ordering (seg[N] $\to$ segment $N{+}1$ assumed) and
        digit-select polarity.
  \item \textbf{Fix \texttt{tools/env.sh} under zsh} --- its
        \texttt{BASH\_SOURCE} fallback points at a stale path from an
        old machine; as a workaround the oss-cad-suite environment was
        sourced directly.
\end{itemize}

\section{Files in \texttt{projects/final\_8/}}
\begin{tabular}{@{}lll@{}}
\toprule
File & Status & Notes \\
\midrule
\texttt{top.v} & one syntax fix left & written by author; line-2 commas pending \\
\texttt{tb\_top.v} & new this session & 16-case sweep + VCD dump \\
\texttt{arty.xdc} & rewritten & Pmod SSD on JA; verified through P\&R \\
\texttt{build\_and\_flash.sh} & existing & end-to-end build script \\
\bottomrule
\end{tabular}

\end{document}
